Coherent IR-Hyper-Raman Four Wave Mixing Vibrational Spectroscopy

### 2024-07-28 paper.tex
following up to ddk edits
Added new time ordering section w better figure (I hope)
Adjusted text based upon JCW and DDK edits
More on slow convergence of 1/w-w' as w->0
Attempted to clean other things up

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics in isotropic media.
% Three wave mixing vibrational spectroscopies, such as vibrational sum frequency generation (vSFG) are similarly used to probe the spectroscopy and dynamics of interfacial species.
Most of these methods rely on infrared and/or Raman transitions, but methods involving hyper-Raman transitions are also possible. 
% Implementing a nonlinear spectroscopy involving hyper-Raman transitions can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (Hyper-DFG, or HDFG) spectroscopy is an underdeveloped four wave mixing vibrational spectroscopy based on the hyper-Raman transition. 
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored.  % DDK: what aspects do we explore?
To this end, we investigate the selection rules and behavior of HDFG spectroscopy.
% We explore the selection rules of singly and doubly resonant HDFG spectroscopy through a Placzek and vibronic picture and show HDFG is a hybrid IR-hyper-Raman technique.
% Since all infrared active modes have non-zero hyper-Raman activity, this makes HDFG a type of upconverted infrared spectroscopy.
% Through a simple comparison test, we demonstrate that HDFG output is shown to be similar in magnitude to vibrational sum frequency generation (vSFG) in a transmission geometry.
% Since HDFG can also be implemented as a two beam experiment in a geometry similar to vSFG, HDFG is a feasible third order spectroscopy for vSFG practitioners.
% Through a simple treatment of orientational averaging, HDFG provides a simple method to extract hyper-Raman hyperpolarizabilities ($\beta$) of infrared active vibrations.
HDFG shows promise as a method to disentangle vibrational spectra and dynamics in isotropic systems.

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325}
% DDK: quickly get to vibrational spectroscopy and vibrational-electronic coupling
To target vibrational and electronic coupling, methods commonly make use of Raman transitions (DOVE, TSF, CARS, SRS, RISRS, and citations).
Here we explore the less-utilized hyper-Raman transitions in CMDS four wave mixing spectroscopy.

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a+\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$.
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through an inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been utilized in several CMDS studies.\cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Bonn2024, McDonnell2024,Wang2021}

Hyper-Raman difference frequency generation (HDFG) spectroscopy and hyper-Raman sum frequency generation (HSFG) spectroscopy are four wave mixing CMDS methods based on direct excitation of a vibrational mode through IR absorption and subsequent scattering to or from the vibrational mode through a hyper-Raman transition.
In HDFG/HSFG, an infrared pulse is resonant with a vibrational mode, while two other excitations stimulate the hyper-Raman scattering process.
The infrared excitation frequency selects the vibrational modes to be stimulated through the hyper-Raman transition.
The other two excitation frequencies control electronic enhancement for the scattering event.

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares it with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require non-centrosymmetry.
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to FWM cascades (CITE).
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just the vibrational mode and the electronic state (``evx'' coupling). 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as 2-color experiments (with $\omega_2=\omega_3$); vSFG or vDFG setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

@@

% DDK: Quick placeholder caption.  Needs to be refined.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{taxonomy.png}
	\caption{
		Spectroscopic methods for investigating vibrational-electronic coupling.
		The interactions of light with matter are shown with Wave Mixing Energy Level (WMEL) diagrams.\cite{RN286}
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
	}
	\label{fig:comparisonwmel}
\end{figure*}

Inspired by recent work that demonstrated the presence of HDFG in ultrafast DOVE and terahertz-IR-visible (TIRV) experiments, we investigate the parameters which result in HDFG output. \cite{Cho2000, Bonn2024, McDonnell2024}
This paper is organized as follows.
In \autoref{steadystate}, we first identify the selection rules that drives HDFG in the absence of electronic resonance. 
The selection rules of electronically-resonant HDFG are then identified through the Herzberg-Teller dipole expansion.
Not surprisingly, we find the selection rules are nearly identical to those of hyper-Raman scattering.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, the site selective properties and useful experimental aspects of HDFG are discussed in \autoref{quant}.
A new method for extracting hyper-Raman hyperpolarizabilities ($\beta_{ijk}$) from HDFG spectra is identified and discussed.

@@

% DDK: compiling caveats/scope narrowing statements throughout the paper; to be refined
% Singly resonant hyper difference generation spectroscopy (HDFG) spectroscopy has potential to provide deeper insight into single quantum decoherence times and molecular orientation in condensed systems.
% Similarities between HDFG spectroscopy, infrared spectroscopy and spontaneous hyper-Raman scattering have been noted previously. \cite{RN352, Bonn2024, McDonnell2024}
In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the information content is similar.
We note that HDFG emission can be phase-matched in media with normal dispersion,\cite{RN278} while HSFG cannot.

Before presenting selection rules, we note that a potential interferant in HDFG spectroscopy is a sequential cascade of second order DFG and SFG processes.\cite{RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or noncentrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume negligible second order cascades. % , and use the pathways drawn in \autoref{fig:hdfg} to examine HDFG response.

It is useful to expose relationships between transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004}
Under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation): \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF(\omega_4) \langle \gamma_{ijkl} \rangle
\end{equation}
where $N$ is a number density, $F$ is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
The brackets indicate an orientational (or rotational) average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.

% To make the connection between HDFG and hyper-Raman scattering, we investigate its gross selection rules.
% DDK: note that with the current placeholder figure, the state indices will have to be changed in our expressions for e.g. mu, Delta
By propagating density matrix elements in the steady state limit under the rotating wave approximation, the HDFG hyperpolarizability is \cite{RN133}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl}^{vg} =	- \sum_{m, n} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{vn}_{i} \mu^{nm}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{nv} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
$D$, the Maker-Terhune degeneracy factor, accounts for permutation symmetry of the excitation fields.\cite{RN134} 
For a HDFG experiment using two (three) distinct input fields, $D = 3 (6)$.

% DDK: or should this be "Vibrational Selection Rules"?  Does Placzek == nonresonant?
\subsection{Electronically Non-resonant: Vibrational Selection Rules}
When $\omega_2+\omega_3$ is significantly detuned from resonance,\cite{Placzek1934, Long1970, Altmann1982} \autoref{sivegamma} 
% DDK: broken thought here
When there is no electronic resonance, HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN352}
We first investigate the HDFG selection rules through a Placzek type formalism.
Contracting over the virtual states forms the hyper-Raman hyperpolarizability $\beta$:\cite{Long1970} 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}.
\end{equation}
All infrared active transitions are hyper-Raman active, making HDFG allowed for any infrared active transition. \cite{Andrews1978}
This selection rule is generally valid for any HDFG when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.
% It is important to note that the symmetry properties of $\beta_{ijk}$ change dependent upon whether $\omega_2$ and $\omega_3$ are degenerate. \cite{Denisov1986, Kozich2007}  
% DDK: not important, I think?  does this have any bearing on our analysis?

The formula gets simpler when we consider IR excitation of a normal mode.  
Assume state $v$ is the $n^{\text{th}}$ reduced normal mode.
We Taylor expand the dipole and first hyperpolarizability operators to first order in the reduced normal mode coordinate $Q_n$ about equilibrium:\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_n}\right)_0 Q_n 
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_n}\right)_0 Q_n
	\end{equation}
\end{subequations}
Substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl}^{vg} =	-\frac{1}{\varepsilon_0} \frac{1}{8D m_n \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_n}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_n}}\right)_0  \rho_{gg}
\end{equation}
% DDK: do we call m_n the reduced mass, or the effective mass?
where we have used $\mel{v}{Q_n}{g} = \sqrt{\frac{\hbar}{2m_n\omega_{ng}}}$, $m_n$ is the reduced mass of the normal mode, and $\omega_{ng}$ is the characteristic frequency.\cite{RN230}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for normal modes in the harmonic limit.

% DDK: wait, this is a little different; we are considering w2 as resonant with a vibrational state; we have not considered that yet...
% leaving this out for now
% When electronically non-resonant, there are additional restrictions on the order of the excitation pulses.
% Specifically, when pulse 2 or 3 interacts before pulse 1, and pulse 2 is resonant with a vibrational state, there is interference between competing pathways that is highly destructive.  
% \autoref{fig:hdfg} illustrates the competing pathways; (a) and (b) ((c) and (d)) have the same pulse ordering, but in one pathway pulse 1 stimulates a ket-side transition and in the other a bra-side transition, which imparts a $\pi$ phase shift difference.(CITE)

@@

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model. \cite{Kundu2022}
Using a two-well system described by 
\begin{subequations}
	\begin{equation}
		H_g = \frac{p^2}{2m} + \frac{1}{2} \hbar \omega q^2
	\end{equation}
and
	\begin{equation}
		H_e = \frac{p^2}{2m} + \frac{1}{2} \hbar \omega (q-\xi)^2 +\hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ is the momentum of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi$ (\autoref{fig:doubres_spec}a). \cite{Carlson1988thesis} 
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
The spectrum is simulated for an electronic state which is both one and two photon allowed.
A simulated spectrum (\autoref{fig:doubres_spec}b) highlighting the contributions of these terms on potential wells with $\xi = 0.5$ dissects the $A$ and $B$ contributions to the HDFG spectrum.
This stimulated hyper-Raman type spectrum would be the expected result in a HDFG experiment when investigating the vibronic structure of the electronic state coupled to $\ket{g,v}$.

>>>>

\begin{abstract}
Nonlinear, four wave mixing vibrational spectroscopies are often used to probe intramolecular vibrational coupling and relaxation dynamics in isotropic media.
Most of these methods rely on infrared and/or Raman transitions, such as sum-frequency generation (SFG) spectroscopy, but methods involving hyper-Raman transitions are also possible. 
The additional number of tensor elements in the hyper-Raman hyperpolarizability have been useful for probing condensed phase media. 
Implementing a nonlinear spectroscopy involving hyper-Raman transitions can provide a useful analogue to infrared and Raman based nonlinear spectroscopies to understand the spectroscopy and dynamics of isotropic systems.
Hyper Difference Frequency Generation (Hyper-DFG, or HDFG) spectroscopy is an underdeveloped four wave mixing vibrational spectroscopy based on the hyper-Raman transition. 
Despite several experimental reports on HDFG, its spectroscopic properties have not been fully explored.  
To this end, we investigate the selection rules and behavior of HDFG spectroscopy as an upconverted infrared spectroscopy and as a probe of vibronic coupling in molecular systems.
HDFG shows promise as a method to disentangle vibrational spectra and dynamics in isotropic systems.

@@

\section{Introduction}
Coherent multidimensional spectroscopy (CMDS) is a family of nonlinear spectroscopy methods that form the optical analogue of multidimensional nuclear magnetic resonance (NMR) spectroscopy.\cite{Cho2008}
Multiresonant, four wave mixing CMDS experiments, first proposed by Oudar and Shen in 1980,\cite{RN307} directly probe coupling and correlations between different vibrational, electronic, and vibronic states. \cite{RN281, RN103, Cho2008} 
Four wave mixing CMDS has resolved anharmonicities, ultrafast dynamics, and other inter- and intra-molecular couplings in numerous systems. \cite{Cho2008, Gaynor2017, Ziegler2018, Ogilvie2019, Bonn2021, RN325}
To target vibrational and electronic coupling, methods commonly make use of Raman transitions, e.g., doubly vibrationally enhanced spectroscopy (DOVE), triple sum frequency (TSF), coherent anti-Stokes Raman spectroscopy (CARS), resonant impulse stimulated Raman spectroscopy (RISRS).\cite{RN103, Dhar1994, RN345, Cho2000, RN491}
Here we explore the less-utilized hyper-Raman transitions in four wave mixing CMDS.

\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66in]{taxonomy.png}
	\caption{
		Some spectroscopic methods for investigating (top) electronic-vibrational coupling and (bottom) e-v-x (x: e, v, e-v, e-v-e) coupling through different order nonlinear processes.
		Note that the coherent $\chi^{(5)}$ e-v-x processes are difficult to isolate due to $\chi^{(3)}$ cascades. 
		The light-matter interacitons are depicted using Wave Mixing Energy Level (WMEL) diagrams.\cite{RN286}
		Solid and dashed horizontal lines indicate real and virtual states, whereas solid and dotted arrows indicate ket and bra side transitions, respectively. 
	}
	\label{fig:comparisonwmel}
\end{figure*}

The hyper-Raman transition is the nonlinear, two photon analogue to the Raman transition.\cite{Terhune1965, Cyvin1965, Andrews1978}
Two photons of frequency $\omega_a+\omega_b$ inelastically scatter with matter, promoting or demoting a vibrational mode of frequency $\omega_v$ and emitting a single photon of frequency $\omega_a + \omega_b \mp \omega_v$. 
The single photon emission is a significantly different frequency from the excitation frequencies, making rejection of excitation scatter much easier.
Unlike Raman, all infrared active modes are hyper-Raman active, which allows infrared active modes to be probed through an inelastic scattering event. \cite{Andrews1978}
One drawback is that hyper-Raman transitions are typically weak compared to their Raman counterparts, even when accounting for the high intensities available to ultrafast pulsed lasers.\cite{RN515, Kelley2010}
CMDS methods partially mitigate this issue because emission is spatially coherent so that all emission is directional and easily collected.
In fact, hyper-Raman transitions have been observed in several CMDS studies.\cite{Zilian1994, RN350, RN416, RN351, RN352, RN353, Chen1998, RN362, RN418, Wang2021, Bonn2024, McDonnell2024}

Hyper-Raman difference frequency generation (HDFG) spectroscopy and hyper-Raman sum frequency generation (HSFG) spectroscopy are four wave mixing CMDS methods based on direct excitation of a vibrational mode through IR absorption and subsequent scattering to or from the vibrational mode through a hyper-Raman transition.
In HDFG/HSFG, an infrared pulse is resonant with a vibrational mode, while two other excitations stimulate the hyper-Raman scattering process.
The infrared excitation frequency selects the vibrational modes to be stimulated through the hyper-Raman transition.
The other two excitation frequencies control electronic enhancement for the hyper-Raman scattering event.

\autoref{fig:comparisonwmel} shows both HDFG and HSFG processes and compares them with related spectroscopies that investigate vibrational-electronic coupling.
The three-wave mixing ($\chi^{(2)}$) techniques are often less applicable as they require non-centrosymmetry.\cite{RN227}
Hyper-Raman scattering itself is a six-wave mixing ($\chi^{(5)}$) technique; stimulated hyper-Raman will be weak and susceptible to FWM cascades.\cite{RN515, RN243, Cho2000_Cascade}
A coherent six-wave mixing analogue of hyper-Raman scattering, coherent anti-Stokes hyper-Raman spectroscopy (CAHRS), has been proposed and explored theoretically by several authors; however, it is also highly susceptible to FWM cascades.\cite{Berger1978, Bjarnason1980, RN243, Cho1997, Cho1998, Cho2000_Cascade}
Of the many four-wave mixing techniques, HDFG and HSFG have strong similarity to Raman, stimulated Raman, and RISRS in that the electronic state must remove or add a $v$ quantum to the ground state (``ev'' coupling).
With the other four-wave mixing techniques (\autoref{fig:comparisonwmel}, second row), the output signal depends on more states than just a single vibrational mode and electronic state (``evx'' coupling). 
As a result, the relationship between the output and vibrational-electronic coupling is more detailed, but also more complicated.
HDFG, HSFG and Raman are more direct probes of electronic-vibrational coupling.
Notably, HDFG and HSFG can be implemented as 2-color experiments (with $\omega_2=\omega_3$); vibrational sum frequency generation (vSFG) or difference frequency generation (vDFG) setups can be trivially reconfigured to perform HSFG or HDFG measurements, respectively, to investigate ev coupling in bulk systems.

@@

Inspired by recent work that demonstrated the presence of HDFG in ultrafast DOVE and terahertz-IR-visible (TIRV) experiments, we investigate the parameters which result in HDFG output. \cite{Cho2000, Bonn2024, McDonnell2024}
This paper is organized as follows.
In \autoref{steadystate}, we first identify the selection rules that drives HDFG in the absence of electronic resonance. 
The selection rules of electronically-resonant HDFG are then identified through a Herzberg-Teller dipole expansion.
HDFG is found to be allowed in any harmonic system that has infrared active vibrations.
A simple harmonic oscillator model system is used to simulate HDFG spectra and illustrate how the excited state potential energy surface affects the hyper-Raman excitation spectrum.
After developing selection rules, the site selective properties, useful experimental aspects of HDFG, and a new scheme to extract the hyper-Raman hyperpolarizability ($\beta_{ijk}$) from HDFG spectra are discussed in \autoref{quant}.
The main findings are summarized in \autoref{conclusion}.

@@

In this section, we investigate the properties of HDFG and make the connections between HDFG and hyper-Raman scattering explicit.
HSFG will not be discussed, but the application of our treatment to HSFG is straightforward and the informational content is similar.
We note that HDFG emission can be phase-matched in media with normal dispersion, while HSFG cannot.\cite{RN278}

Before presenting selection rules, we note that a potential interferant in HDFG spectroscopy are direct and sequential cascades of second order vDFG and vSFG processes.\cite{RN297, RN302, RN301}
While unimportant in achiral isotropic systems,\cite{Belkin2000} second order cascades may become important in media where SFG and DFG are allowed, e.g. chiral media, interfaces or noncentrosymmetric media. 
Nevertheless, in the discussion that follows for the rest of this manuscript, we will assume negligible second order cascades.

It is useful to expose relationships between transition dipoles and hyper-Raman hyperpolarizabilities in the driven limit. \cite{Simpson2004}
Under the electric dipole approximation, the I$^{th}$ component of the third order nonlinear output polarization, ${P}^{(3)}_I$, of a four wave mixing process, induced by electric fields $E_J$, $E_K$, and $E_L$ at output frequency $\omega_4=-\omega_1 + \omega_2 + \omega_3$ is written as (using Einstein summation) \cite{RN307}
\begin{equation} \label{polarization}
{P}^{(3)}_I (\omega_4)  = \chi^{(3)}_{IJKL} E_J(\omega_3) E_K(\omega_2) E_L(\omega_1) 
\end{equation}
where $\chi^{(3)}_{IJKL}$ is the IJKL element of the third order electrical susceptibility, a rank four tensor, generally written as
\begin{equation}
	\chi^{(3)}_{IJKL} = NF(\omega_4) \langle \gamma_{ijkl} \rangle
\end{equation}
where $N$ is a number density, $F$ is the Lorentz local field factor, and $\gamma_{ijkl}$ is the third order polarizability (i.e., second hyperpolarizability). 
The brackets indicate an orientational (or rotational) average.\cite{Andrews1977}
Uppercase letters refer to laboratory frame coordinates and lower case letters refer to corresponding molecular frame coordinates.

By propagating density matrix elements in the steady state limit under the rotating wave approximation, the HDFG hyperpolarizability is \cite{RN133}
\begin{equation}\label{sivegamma}
		\gamma_{ijkl}^{vg} =	- \sum_{e,m} \frac{1}{\varepsilon_0} \frac{1}{4D} \frac{1}{\hbar^3} \frac{\mu^{ve}_{i} \mu^{em}_{j} \mu^{mg}_{k} \mu^{gv}_{l} }{\Delta_{ev} \Delta_{mv}\Delta_{gv}}  \rho_{gg}
\end{equation}
where: $\mu^{ab}_{j}$ is the $j^{th}$ element of $\mel{a}{\vec{\mu}}{b}$, $\Delta_{kl} = \omega_{kl} - \omega_{j} - i\Gamma_{kl}$, $\omega_j$ is the frequency of the j$^{th}$ input field, $\Gamma_{kl}$ is the dephasing of $\rho_{kl}$ and $\rho_{gg}$ is the ground state population.
$D$, the Maker-Terhune degeneracy factor, accounts for permutation symmetry of the excitation fields.\cite{RN134} 
For a HDFG experiment using two (three) distinct, non-degenerate input fields, $D = 3 (6)$.

% DDK: or should this be "Vibrational Selection Rules"?  Does Placzek == nonresonant?
\subsection{Vibrational Selection Rules}

When there is no electronic resonance, HDFG has also been referred to as singly vibrationally enhanced (SIVE) spectroscopy. \cite{RN352}
It is useful to investigate the selection rules without vibronic resonance to understand how HDFG can be used as an upconverted infrared spectroscopy.
We first investigate the HDFG selection rules through a Placzek type formalism.
When $\omega_2+\omega_3$ is significantly detuned from resonance and the vibronic nature of the virtual states is ignored, the Placzek approximation can be used to study selection rules in terms of the hyper-Raman hyperpolarizability, $\beta$. \cite{Placzek1934, Long1970, Altmann1982}
Summing over the virtual states $\ket{e}$ and $\ket{m}$ forms the hyper-Raman hyperpolarizability so that\cite{Long1970} 
\begin{equation}\label{sivebeta}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{4D \hbar}\frac{\beta^{vg}_{ijk} \mu^{gv}_{l}}{\Delta_{gv}} \rho_{gg}.
\end{equation}
Since all infrared active transitions are hyper-Raman active, HDFG is allowed for any infrared active transition. \cite{Andrews1978}
This selection rule is generally valid for any HDFG or HSFG process when $\omega_2$ and $\omega_3$ are sufficiently detuned from any resonance.

The general selection rule can be simplified by considering excitation of an infrared active normal mode.  
Assume state $v$ is the $v^{\text{th}}$ normal mode.
We Taylor expand the dipole and first hyperpolarizability operators to first order in the normal mode coordinate $Q_v$ about equilibrium:\cite{Long1970, Shen90}
\begin{subequations}
	\begin{equation}
		\mu_l = \mu_{l,0} + \left(\frac{\partial \mu_l}{\partial Q_v}\right)_0 Q_v 
	\end{equation}
	\begin{equation}
		\beta_{ijk} = \beta_{ijk,0} + \left(\frac{\partial \beta_{ijk}}{\partial Q_v}\right)_0 Q_v
	\end{equation}
\end{subequations}
Substituting into \autoref{sivebeta} gives the HDFG hyperpolarizability to $\order{Q_n}$ as \begin{equation}\label{SIVEselection}
	\gamma_{ijkl} =	-\frac{1}{\varepsilon_0} \frac{1}{8D \omega_{vg}}  \frac{1}{{\Delta_{gv}}} \ \left(\frac{\partial \beta^{vg}_{ijk}}{\partial Q_v}\right)_0 \left({\frac{\partial \mu^{gv}_{l}}{\partial Q_v}}\right)_0  \rho_{gg}
\end{equation}
where we have used $\mel{v}{Q_v}{g} = \sqrt{\frac{\hbar}{2\omega_{vg}}}$ (in the limit $v-g = 1$) and $\omega_{vg}$ is the characteristic frequency.\cite{RN459}
Since this expression is non-zero in the harmonic oscillator limit, HDFG output is allowed for normal modes in the harmonic limit.

\subsection{Time Ordering}
It is useful to briefly discuss the impact of time ordering on HDFG output.
Dependent upon the time-ordering schemes, it is possible to eliminate output through destructive interference.
While we have only isolated one specific HDFG process (\autoref{fig:comparisonwmel}), there are at least fourteen different WMELs which can provide HDFG output in the driven limit. \cite{RN352}
If $\omega_2$ ($\omega_1$) becomes resonant (non-resonant), where $\omega_2 > \omega_1$, then other HDFG pathways appear (\autoref{fig:sivewmel2}).\cite{McDonnell2024} 
While the methods in \autoref{fig:sivewmel2} possess identical selection rules to \autoref{fig:comparisonwmel}, the WMEL diagrams presented in \autoref{fig:sivewmel2} will interfere, as they share the same phasematching condition but have oppositely signed amplitudes.
\begin{figure*}[!htbp]
	\centering
	\includegraphics[width=6.66 in]{figures/timeorderedwmel.png}
	\caption{WMEL diagrams of HDFG pathways for when (a,b) $\omega_1$ (black arrow) interacts first and (c-f) $\omega_2$ (orange arrow) interacts first. 
		The states are labeled as $\ket{a,b}$, where $a$ refers to an electronic state and $b$ refers to quanta in a vibration on $a$.
		In all cases, $\omega_2 > \omega_1$.
	}
	\label{fig:sivewmel2}
\end{figure*}

In the limit where $\omega_2+\omega_3$ are detuned from an electronic resonance, it becomes simple to interpret the interference schemes. 
The case where $\omega_1$ is resonant and $\omega_2 + \omega_3$ are nonresonant yields two contributed diagrams, assuming $\omega_1$ interacts with the system first.
These WMEL diagrams possess identical phase and transition moments, which yields total constructive interference.
Conversely, diagrams (c,d) and (e,f), when detuned from an electronic resonance, have identical transition moments but opposite phase. 
As a result, the output from (c,d) and (e,f) completely destructively interfere; no signal can be obtained from these pathways when significantly detuned. \cite{RN287, McDonnell2024}

When $\omega_2 + \omega_3$ become electronically resonant, the interference changes for the \autoref{fig:sivewmel2}c-f pathways.
Note that only pathways (c,e,f) possess identical transition moments.
As a result, pathways (c,d) survive and can provide output when $\omega_2+\omega_3$ is resonant with an electronic state. 
These results provide a simple rule of thumb for collecting HDFG spectra: use \autoref{fig:sivewmel2}a,b pathways to measure spectra when significantly detuned from an electronic resonance, and any pathway can measure spectra when electronically resonant.
However, since electronic states commonly have short dephasing times and have significantly inhomogeneous broadening,\cite{Dong2015, Lewis2015} it is likely the \autoref{fig:sivewmel2}f pathway will become triply resonant and possibly contribute greater than the other pathways. 
Such instances are system dependent. 
For consistency, discussion will be isolated to the \autoref{fig:sivewmel2}a pathway in the sections that follow.

@@

The impact of vibronic coupling in the 2D spectrum is made clear by using a simple model. \cite{Kundu2022}
Using a two-well system described by 
\begin{subequations}
	\begin{equation}
		H_g = \frac{p^2}{2m} + \frac{1}{2} \hbar \omega q^2
	\end{equation}
and
	\begin{equation}
		H_e = \frac{p^2}{2m} + \frac{1}{2} \hbar \omega (q-\xi)^2 +\hbar \omega_{eg}
	\end{equation}
\end{subequations}
for the ground ($H_g$) and first excited ($H_e$) states, where $p$ is the momentum of the normal mode, the A and B contributions can be evaluated using Franck-Condon and Herzberg-Teller integrals tabulated in terms of the dimensionless normal mode $q$ and offset parameter $\xi$ (\autoref{fig:doubres_spec}a). \cite{Carlson1988thesis} 
Note that $q = \sqrt{\frac{m^2\omega}{\hbar}} Q$.
The spectrum is simulated for an electronic state which is both one and two photon allowed.
A simulated spectrum (\autoref{fig:doubres_spec}b) highlighting the contributions of these terms on potential wells with $\xi = 0.5$ dissects the $A$ and $B$ contributions to the HDFG spectrum.
This stimulated hyper-Raman type spectrum would be the expected result in a HDFG experiment when investigating the vibronic structure of the electronic state coupled to $\ket{g,1}$.
